---
title: "Introduction to pdfr"
author: "Stephanie J. Spielman"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

The primary of the `pdfr` package is to save you, the time-limited user, from having to type 12 (yes, 12!) extra characters when printing out full data frames. Intruiged? Read on...

This mostly comes in handy when trying to view the *entire* data frame (or more likely, tibble) being piped through a `tidyverse` pipeline. One of the generally intelligent decisions that `tibble` makes is to only print the first several columns, and first several rows, to save your eyeballs from scrolling:

```{r message=FALSE, warning=FALSE}
library(ggplot2) #> contains example data set
library(dplyr)   #> allows demo of piping
library(tibble)  #> show demo with tibbles

msleep.tib <- as.tibble(msleep)

msleep.tib %>% filter(vore == "omni")
```

In this case, however, you can't actually see all the columns. Sometimes, it is actually desirable to view the entire output. 

On one hand, this sort of problem can be solved by setting tibble options, e.g. `options(tibble.print_max = Inf)`. But maybe you can't remember to type this line, or maybe you only *sometimes* want to print the whole data.

In those circumstances, you'll often find yourself piping into `print.data.frame()`:

```{r}
msleep.tib %>% filter(vore == "omni") %>% print.data.frame()
```


That's all well and good, but *16 characters*?! How can lazy programmers like you and me cope with this undue burden? Finally, the answer has arrived - you will find salvation in the `pdfr()` package:
```{r}
library(pdfr)

#> Hurray, only 4 characters!
msleep.tib %>% filter(vore == "omni") %>% pdfr()
```

Now that felt much better. Phew.

<br><br>
There are also two other useful functions, `hpdfr()` and `tpdfr()` which will print the head or tail, respectively, of the full data frame. Below are examples of using `hpdfr()`, and `tpdfr()` works in the same way.
```{r}
#> Print the full head
msleep.tib %>% filter(vore == "omni") %>% hpdfr()


#> Print first 2 rows with the extra argument to hpdfr()
msleep.tib %>% filter(vore == "omni") %>% hpdfr(2)
```




