---
title: "Introduction to pdfr"
author: "Stephanie J. Spielman"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

The primary of the `pdfr` package is to save you, the time-limited user, from having to type 12 (yes, 12!) extra characters when printing out full data frames. We introduce the function `pdfr()` as a simple wrapper for `print.data.frame()`. 

This mostly comes in handy when trying to view the *entire* data frame (or more likely, tibble) being piped through a `tidyverse` pipeline. One of the generally intelligent decisions that `tibble` makes is to only print the first several columns, and first several rows, to save your eyeballs from scrolling:

```{r message=FALSE, warning=FALSE}
library(ggplot2) #> contains example data set
library(dplyr)   #> allows demo of piping
library(tibble)  #> show demo with tibbles

msleep.tib <- as.tibble(msleep)

msleep.tib %>% filter(vore == "omni")
```

In this case, however, you can't actually see all the columns.
Sometimes, it is actually desirable to view the entire output. Normally you'd have to type `print.data.frame()`:

```{r}
#> So many characters to type!
msleep.tib %>% filter(vore == "omni") %>% print.data.frame()
```

But now, you can do it in 4 instead of 16 characters! What a relief!
```{r}
library(pdfr)

#> Hurray, only 4 characters!
msleep.tib %>% filter(vore == "omni") %>% pdfr()
```

There are also two other useful functions, `hpdfr()` and `tpdfr()` which will print the head or tail, respectively, of the full data frame. Below are examples of using `hpdfr()`, and `tpdfr()` works in the same way.
```{r}
#> Print the full head
msleep.tib %>% filter(vore == "omni") %>% hpdfr()


#> Print first 2 rows with the extra argument to hpdfr()
msleep.tib %>% filter(vore == "omni") %>% hpdfr(2)
```




